Are populations like a circuit? The relationship between isolation by distance and isolation by resistance

### 2019-01-05 review-responses.tex

%%%%%%
%%
%%  Don't reorder the reviewer points; that'll mess up the automatic referencing!
%%
%%%%%

\begin{minipage}[b]{2.5in}
  Resubmission Cover Letter \\
  {\it Molecular Ecology Resources}
\end{minipage}
\hfill
\begin{minipage}[b]{2.5in}
    Erik Lundgren \\
    \emph{and} Peter Ralph \\
  \today
\end{minipage}

\vskip 2em

\noindent
{\bf To the Editor(s) -- }

\vskip 1em

We are writing to submit a revised version of our manuscript, 
``Are populations like a circuit? The relationship between isolation by distance and isolation by resistance''.

...

\vspace{5em}

\noindent \hspace{4em}
\begin{minipage}{3in}
\noindent
{\bf Sincerely,}

\vskip 2em

{\bf 
Erik Lundgren and
Peter Ralph
}\\
\end{minipage}

\vskip 4em

\pagebreak
\setcounter{page}{1}

%%%%%%%%%%%%%%
\reviewersection{AE}

\begin{quote}
We received five reviews for your submitted article. While three are quite positive, two others are more critical. In the balance of comments, all reviewers felt that your article should be published in MER, though significant improvements were requested. Therefore, my decision is to request a revision before accepting your article.

Please respond to all major comments raised by the reviewers. In particular, try to address the following points: 1) Revise figures for clarity and legibility. 2) Include an empirical dataset, as suggested by REV 2. 3) Rewrite for a broader audience of biologists (translating some difficult concepts when possible and as appropriate). 4) Clarify aspects of the statistical methods as indicated by several reviewers.
\end{quote}

%%%%%%%%%%%%%%
\reviewersection{1}

\begin{quote}
    The authors present a new Bayesian method to fit movement rates between a
    set of demes to pairwise genetic differentiation, utilizing a MCMC. They first
    state the basic equations, which closely resemble existing theory of stepping
    stone models. They extensively compare their computational method to resistance
    based methods using an own implementation of the latter, both "under the model"
    as well as using various quite realistic forward in time simulations.
    Resistance based methods are much faster and can handle much more demes than
    the method presented here (no case with more than a 5x5 grid is shown).
    Therefore, as the authors point out, this work will likely not contest the use
    of resistance distances for visualizing spatial genetic patterns (which is
    often done in an exploratory way). But importantly, the method here can infer
    asymmetric and biased rates, unlike almost all existing methods, which presents
    a major advance -- and certainly fills a niche.  
\end{quote}

\begin{quote}
    Despite the computational limitations, I believe that this paper will provide
    researchers a good starting point for future development, by both giving a
    review about difficulties and also a first idea and example for an
    implementation.

    Overall, I think this article makes a very valid contribution for Molecular
    Ecology Ressources. The authors put their work into the right context and
    show good knowledge of the relevant literature. The writing is both clear
    and concise. Importantly, they do not shy away from pointing out the limits
    of their method, and the very stringent (often unrealistic) assumptions
    such methods usually make.

    I have several comments, but none of them are ``make-or-break'' or fatal flaws.
    The manuscript is already in a good state. But I hope the authors can use my
    comments to further improve their work.  
\end{quote}

\begin{point}{}
    Clearer point out the current computational limitations of your coalescence
time method. It seems it gets tricky with more than about 30 demes? What is the
run-time then, does the whole thing blow up? Give an idea how big is too big!
\end{point}

\reply{
}

\begin{point}{}
    What is the exact scaling? The bottleneck seems to be calculating the
    coalescence times with equation (7). You only say that it ``can be solved
    with standard linear algebra''. Are you doing Gaussian elimination or
    something else worse than $N^2$? Do you use any speed-ups to
    deal with the sparse nature of G for neighbor-only migration? Is it faster
    than getting the coalescence times by iterating time-discretized
    recursions?  It would be helpful for further progress to describe the exact
    scaling and the exact algorithm you use.
\end{point}

\reply{
}

\begin{point}{}
    The title of the manuscript is a bit misleading. One would not guess that
    the introduction of a method to infer biased and asymmetric gene flow is the
    main point.  ``Isolation by distance'' does not appear often in the manuscript,
    neither its classic literature (Wright, Maryuama, Malecot, Barton etc.) nor
    direct investigation of genetic differentiation with geographic distance. But
    these things are also not the scope of this paper.  
\end{point}

% Citation:
% The Effects of Genetic and Geographic Structure on Neutral Variation
% Brian Charlesworth, Deborah Charlesworth and Nicholas H. Barton
% Annual Review of Ecology, Evolution, and Systematics
% Vol. 34 (2003), pp. 99-125

\reply{
}

\begin{point}{}
    You could write an outlook stating that your method could be used to fit
    parameterized models with few parameters. For instance, in case the putative
    location of a barrier or an area of biased gene flow is known, one could try to
    infer it without fitting ALL movement parameters, but a simple parametric
    model, which would implicitly take care of the regularization.  
\end{point}

\reply{
}

\begin{point}{}
    You use mean absolute error in movement rate to describe the model fit. I
    believe this measure can be quite misleading for exponential movement rates.
    Large rates can be notoriously hard to infer (are you super or super-super well
    connected is almost the same), so it is easy to be off quite a bit, which
    drives the average absolute error in rate. Also in practice, one does not care
    so much about absolute rates but ``relative'' ones. So using a correlation
    measure R**2  of inferred movement rates with actual movement rates would be a
    very useful and natural fit parameter.  
\end{point}

\reply{
}

\begin{point}{}
    You write you use standard methods to check convergence and mixing of the
    MCMC chain (line 182). Given the degeneracy and the high dimensionality,
    where there any problems? E.g., local optima you were stuck sometimes? Or
    was it always pretty straightforward? Maybe write a few sentences on that!
\end{point}

\reply{
}

\begin{point}{\revref}
    $F_ST \neq$ Absolute divergence!!! . $F_ST$ is more a
    co-variance measure, and for instance heavily influence by sampling a batch
    of related individuals (whereas the mean of absolute divergence will be
    unchanged).
\end{point}

\reply{
}

\begin{point}{}
    You have your own implementation of resistance distance methods, by first
    extracting hitting times. However, you allow non-symmetric rates, something
    no existing resistance method would do. Given your observation that there
    is a lot degeneracy for resistance times (adding skew stuff to hitting
    times), this introduces uncertainty that likely cannot be broken such as
    with coalescence methods. This could drive the high uncertainty in Fig 5
    upper right. Given your good results for Fig. 4 for the symmetric case,
    this is a bit surprising. Maybe run one scheme where you FIX to a single
    symmetric movement rate between each pair of demes. This setting would be a
    fairer comparison (and be much closer to what EEMS etc. actually do -- they
    do not even try to infer asymmetric rates).
\end{point}

\reply{
}

\begin{point}{}
    Maybe describe the model you use for you calculation and inference, a
    continuous backward in time Markov model with coalescence rates in demes
    before the actual recursions, and not bit by bit in there.
\end{point}

\reply{
}

\begin{point}{\revref}
    Do you actually use this tensor product for the movement
    rates anywhere? And seems Eq. 7 (which has no Tensor product) is all you
    need.
\end{point}

\reply{
    We included this because we think it's a nice compact way to write, code, 
    and think about the equation,
    and the tensor product equation is actually the one used in the R code.
    It is correct that the equations are equivalent.
}

\begin{point}{Line 115:}
    Not all $ZC$ skew-symmetric would translate to valid Generators 
    $G + Z$. The diagonal elements of G have another condition (they have to be minus
    the sum of all other jumping rates). Would be interesting to explore this a
    bit more -- overall what the dimensionality this degeneracy is (implicit
    function theorem), and also if there are simple toy examples where they pop
    up.
\end{point}

\reply{
}

\begin{point}{Eq. 4:}
    $H$ may not be always invertible ($H$ symmetric and positive off-diagonal
    and 0 on the diagonal is NOT enough). So there might be some pathological
    cases where multiple G link to one $H$, maybe even in some simple toy
    examples.
\end{point}

\reply{
    Good point; we had not justified this.
    Under the standing assumption that the chain is irreducible, 
    $H$ is always invertible.
    We now prove this in the Appendix \llname{rr:H_invertible},
    and have made reference to this fact. \revref
}

\begin{point}{}
    Dimensionality arguments (Line 129 onward) can be a bit misleading when you
    already now the equations have a non-trivial kernel. Set all
    non-neighboring elements of $G$ to 0 might not get rid of parts of the
    kernel. So maybe flag it even more as ``be cautious'' as you already have.
\end{point}

\reply{
}

\begin{point}{}
    You could just call the thing by its name: ``Stepping stone'' model. This
    term has such a rich history in population genetics!
\end{point}

\reply{
}

\begin{point}{}
    In the main text, you never mention that the mean of the exponential
    movement rates for simulations is one, only in a figure. One can also kind
    of see it from the model figures, but it would be nice to have it
    explicitly stated in the main text.
\end{point}

\reply{
}

\begin{point}{}
    In the figures, probably keep model figure and inference results closely
    together. This helps the reader, and makes the figures stand more for
    themselves.
\end{point}

\reply{
}

\begin{point}{}
    Useful result to know: Maryuama’s invariance principle (Nagylaki 1998):
    That the correctly weighted within coalescence time within demes only
    depends on the total population size  (given conservative migration) -- so
    it cannot fluctuate too wildly across demes and movement rates! This fact
    motivates why not fitting too wildly differing within-deme coalescence
    times/diversities.
    % Nagylaki, T., 1998 The expected number of heterozygous sites in a subdivided
    % population. Genetics 149: 1599
\end{point}

\reply{
}

\begin{point}{\revref}
    400 000 generations will look a bit lunatic for someone not in
    popgen who does not know about rescaling. You could give re-scaled, more
    realistic values!
\end{point}

\reply{
}

\begin{point}{Paragraph beginning \revref:}
    Maybe move this section in front of the inference part, as it is useful to
    know the most ``fundamental'' limit of resistance distances.
\end{point}

\reply{
}

\begin{point}{Fig. 7:}
    With EEMS one can see very suspiciously the 4x4 sampling structure - there
    are red dots in almost all 3x3 centers of the sampling squares. Since EEMS
    is quite fast, could you increase the resolution of sampling?
\end{point}

\reply{
}

\begin{point}{}
    You could experiment with visualizations a bit. EEMS got quite popular as
    exploratory visualization tool because it produces visually pleasing plots.
    Do lines with two arrowheads of different sizes near the middle, maybe even
    colored (blue to red) work?
\end{point}

\reply{
}

\reviewersection{2}

\begin{quote}
    This work revisits McRae’s seminal isolation-by-resistance (IBR) model through
    the lens of coalescent theory. The original idea of IBR was to compare patterns of
    genetic distance observed on a landscape with the so-called ``resistance distance''.
    The resistance distance between states i and j in a Markov chain is defined to be
    the ``commute time'': the expected amount of time it takes for the chain to go
    from i to j and back, integrated over all possible paths of the chain. Resistance
    distance comes from electrical circuit theory and is not a model of genetic
    variation across a landscape. Also, it is inherently symmetric in i and j, and
    cannot accomodate asymmetric migration between demes. A more biologically
    realistic model is the structured coalescent. Hence, if data are generated under
    the something resembling the coalescent, then there is scope for improving IBR
    by reducing model misspecification.

    This paper implements such a method and conducts comparative analyses on a
    number of simulated data sets. It seems that in general, this method and the
    resistance-based approaches (as exemplified by EEMS) produce somewhat com-
    parable results so long as migration rates are symmetric. The largest differences
    emerge when gene flow is asymmetric, because this fundamentally violates the
    resistance distance model used by EEMS. Here, the authors show convincingly
    that inferring migration rates using the resistance distance approximation can
    be wildly inaccurate compared to the same estimates obtained using coalescent
    methods. To the extent that gene flow is more likely to be asymmetric in many
    natural populations, this seems like an important finding.

    My overall impression of this work is fairly positive. The paper is very clearly
    written, and the idea of replacing IBR with a method that is more faithful to
    the underlying biology is appealing. The methods section (as well as appendices)
    feature a very nice and readable summary of the mathematical innovations of
    this work, as well as some interesting discussion of issues of identifiability and
    power. Similarly, discussion section in very well done and really strengthens
    the manuscript. It includes a thorough survey of the related work, as well as a
    forthright discussion of advantages and disadvantages of the proposed method,
    and it raises a number of interesting questions and directions for future research.
    Overall, the paper was a pleasure to read.

    I find the main deficiency to be the lack of any analysis of real data. All
    experiments are carried out on simulated data only. This is a puzzling omission.
    The authors went to the trouble of creating an R package (which is good), so
    clearly they intend for the method to be used in practice. Personally, I would
    hesitate to publish the results obtained from a method which had not previously
    been verified to produce reasonable answers in real data. This really feels like
    something the authors ought to have included. It is not necessary that the data
    be novel -- reanalyzing (say) the African elephant dataset in the EEMS paper
    would provide for an interesting comparison (even if it cannot be done at an
    equivalent spatial resolution) and would also go long way towards reassuring
    practitioners.
\end{quote}

\begin{point}{}
    ``Bayesian inference of movement rates'': In their model the authors assume
    that many entries of the migration rate matrix are zero, and place an
    exponential prior on the rest. Is it safe to assume that we will always
    want to constrain rates between discontiguous cells to zero? Also, given
    the valid concerns that they raise elsewhere about identifiability and
    power, I wonder if the authors considered placing a sparse prior on the
    entire G matrix, enabling the model to adapt to the amount of signal while
    shrinking other migration rates to zero. This might have the effect of
    making the resulting non-zero estimates more stable. (I realize that this
    raises potential problems for inference if the underlying graph becomes
    disconnected.)
\end{point}

\reply{
}

\begin{point}{}
    The discussion of ill-conditioning is welcome. From experience, it often
    feels like coalescent-based methods have little power to accurately
    estimate migration unless the rates are large. It might be interesting to
    shift emphasis from estimation to the slightly easier problem of testing:
    how good are these methods at distinguishing between a given null and an
    alternative hypothesis (say, barrier vs. no barrier) for the rate matrix G?
    I think the gain of using a coalescent model might be more apparent in this
    case.
\end{point}

\reply{
}

\begin{point}{}
    This method is computationally limited by equation (7), which has a
    complexity of $O(d^6)$ for the product Markov chain on d demes mentioned a
    few lines below. At least, that is the cost for solving (7) exactly -- but
    given the noisiness of the matrix C, there may be little point in doing
    that anyways. I wonder if the authors have thought at all about obtaining
    an approximate solution to (7), where the error in approximation is chosen
    to in some way to balance out the statistical error inherent in C / D. It
    may also be possible to leverage the fact that the solution G is sparse
    here.  (This is more of a suggestion for future research rather than a
    suggestion for revision.)
\end{point}

\reply{
}

\begin{point}{\revref}
     Do the authors have any other guidance on how to discretize the landscape?
\end{point}

\reply{
}

\begin{point}{}
    Does the issue of having negative entries of G (lines 112-117, Appendix A)
    arise during the course of MCMC sampling, and if so, how is it dealt with?
\end{point}

\reply{
}

\begin{point}{\revref}
     suggest ``commonly'' in place of ``not uncommonly''
\end{point}

\reply{
    Changed.
}

\begin{point}{\revref}
     strongly misleading?
\end{point}

\reply{
    We think this is correct (and reads reasonably well?) as it stands.
}

\reviewersection{3}

\begin{quote}
    This is an interesting paper on a topic that should be of wide interest to
    readers of MER.  The main results are useful: 
    \begin{enumerate}
        \item      landscape genetics models based on resistance theory can be
            misleading if gene flow is asymmetrical
        \item      coalescent-based models deal better with asymmetry but are
            not a panacea for a variety of reasons.
    \end{enumerate}
    On the other hand, I found myself with quite a few questions after reading
    the manuscript.  Raising new questions is not bad in and of itself, but
    most of these were questions that could/should have been clarified in the
    text.  As a consequence, the manuscript is not as effective or useful as it
    might be.  The problem is not in the writing itself, which generally is
    careful and precise.  Instead, the main problem involves translating a very
    technical exercise into something that is accessible to typical readers of
    the journal.  Some examples are described below.

    Although the theme of the paper is relevant to the journal, I think most
    readers of MER would find the manuscript hard going in many places.  For
    example, the text at 102-109 and 136-146 would benefit from translation
    into insights more meaningful to biologists.  The authors do a good job of
    this in discussing results for asymmetric gene flow at lines 296-299.
\end{quote}

\begin{quote}
    The reader also needs to understand the biological consequences of various
    assumptions.  Examples  where this could be improved include:
\end{quote}

\begin{point}{86:}
      if the genome is selectively neutral ...
\end{point}

\reply{
}

\begin{point}{\revref}
     as long as long distance migration is disallowed
\end{point}

\reply{
}

\begin{point}{\revref}
     $C$ is large compared to $\sigma$
\end{point}

\reply{
}

\begin{point}{\revref}
     we can estimate amount of noise well
\end{point}

\reply{
}

\begin{point}{}
    The text at 468ff relating to the assumption of constant pop size and migration
    rates does a good job of explaining these consequences.  Perhaps all
    assumptions could be listed in a single place and their biological
    consequences explained.
\end{point}

\reply{
}

\begin{point}{\revref}
    How exactly is the pairwise genetic distance calculated?  What does ``density
    of sites'' mean?  Is it just the fraction of sites at which two genomes
    differ?  If so it is just pairwise theta as used by Wakeley and others.  
    Eq 1:  what is $R_{ij}$ and how is $q$ calculated?
\end{point}

\reply{
    That's correct; we've reworded this to be more clear.
}

\begin{point}{}
    Eq 1:  what is $R_{ij}$ and how is $q$ calculated?
\end{point}

\reply{
}

\begin{point}{}
    $G$, the movement rate, is clearly a key parameter but is not clearly defined (or
    if it was I missed it).  What are the units of G?  Geographic distance per
    unit time?  What are the time units?  Years?  Generations?  If the latter,
    then does G represent parent-offspring dispersal distance as defined by
    Wright?  Are generations assumed to be discrete?  If not, can a migrant
    leave offspring at various different geographic points?
\end{point}

\reply{
}

\begin{point}{}
    As described, the model deals with groups of individuals (populations) that
    sometimes exchange individuals.  But much of landscape genetics/genomics
    deals with continuously-distributed species that don’t have recognizable
    populations.  Are the results also applicable to scenarios like this?  Why
    or why not?
\end{point}

\reply{
}

### 2019-01-05 isolation_by_coalescence.tex
submission tweaks
review responses

\documentclass{article}
\usepackage{fullpage}
\usepackage{amsmath,amssymb,amsthm}
\usepackage[hidelinks]{hyperref}
\usepackage{graphicx}
\usepackage[utf8]{inputenc}
\usepackage{xcolor}
\usepackage{natbib}
% \usepackage{subcaption}

@@

\newcommand{\R}{\mathbb{R}}
\newcommand{\norm}[1]{\left\lVert#1\right\rVert}
\DeclareMathOperator{\var}{\mathop{\mbox{Var}}}
\DeclareMathOperator{\cov}{\mathop{\mbox{Cov}}}
\DeclareMathOperator{\diag}{\mathop{\mbox{diag}}}

\renewcommand{\P}{\mathbb{P}}
\newcommand{\E}{\mathbb{E}}
\newcommand{\given}{\,\vert\,}
\newcommand{\supp}{\mathop{\mbox{supp}}}
\newcommand{\sgn}{\mathop{\mbox{sgn}}}
\newcommand{\EE}[1]{\E\!\left[#1\right]}
\newcommand{\PP}[1]{\P\!\left\{#1\right\}}
\newcommand{\PPP}[2]{\P_{#1}\!\left\{#2\right\}}
\newcommand{\EEE}[2]{\E_{#1}\!\left[#2\right]}
\newcommand{\EEsup}[2]{\E^{#1}\!\left[#2\right]}

\newcommand{\bone}{\mathbf{1}}

@@

@@

@@

Genetic relatedness is determined by past gene flow,
a product of individual or gamete movements across geographic space.
Genomes therefore retain the traces of this movement,
and can contribute to the inference of how a species moves across a landscape,
which is important for understanding how diseases spread, how species adapt,
and how to retain genetic diversity in threatened species.
This is a potentially important source of information,
as direct observation can be difficult 
(especially of long-distance movement, \citep{cayuelademographic,levin_et_al_seed_dispersal}),
or even impossible, if some of the populations in question no longer exist.

@@

In this paper, we contrast 
coalescence time and resistance distance,
and develop a method to infer movement rates on a discrete landscape
using genetic distances.
As we will see,
although the two methods are conceptually similar,
resistance-based inference may get the wrong answer if data are derived from a coalescent process,
especially in the presence of gene flow asymmetry.
We also explore the question of identifiability.
Resistance-based methods are often used to predict distances based on a few layers of geographical data
(land cover, slope, etcetera).
Some methods find the combination of landscape layers to best explain genetic distance,
which allows very fine geographic resolution \citep{shaffer2017desert}.
Other methods such as EEMS \citep{petkova2016visualizing} infer conductances without such prior information;
however, the geographical resolution is much coarser.
(A recently-released method 
that is based on EEMS but uses haplotype length data and an underlying model of coalescence time
\citep{alasadi2018estimating}
is promising, but still produces similarly coarse maps.)
One reason for this difference in resolution
is that EEMS uses more computationally intensive Bayesian methods.
However, we show that this is a more general problem,
both because of unmodeled ``process noise'' involved in discretization of geographical populations,
and the mathematical structure of the underlying problem.

@@

There are a number of ways to calculate measures of dissimilarity using genetic data.
In this paper we use mean pairwise divergence:
for two genomes, it is calculated as the density of sites at which the two differ.
It is more common in the literature to use $F_{ST}/(1-F_{ST})$,
but the two are highly correlated for good theoretical reasons 
\citep{slatkin1991inbreeding,rousset1997genetic},
especially for SNP data.
The two genomes differ at a site only if there has been a mutation in some ancestor on the lineages leading
from the two genomes back to their most recent common ancestor at that site.
Therefore,
under an infinite-sites model with average mutation rate $\mu$ per site and per meiosis,
pairwise divergence divided by $\mu$
is an unbiased estimate of the mean time to most recent common ancestor,
averaged across the genome \citep{hudson2007variance,ralph2015empirical}.
It is natural to model this process by following the two lineages back through time
(and across space)
until they find their most recent common ancestor,
i.e., until they coalesce.
Suppose we model the geographic distribution of the species
as a collection of randomly mating populations
that exchange occasional migrants,
thus discretizing geographic space.
If the genome is selectively neutral,
then the motion of these lineages back through time
forms a Markov process, the \emph{structured coalescent} \citep{wakeley2005coalescent}.
In this framework, each lineage performs a random walk across the populations,
with movement probabilities depending on the flux of individuals between populations.
This provides the link between genetic distance and random walks.

@@

Suppose instead we are given the hitting times, $H$, and wish to find the movement rates, $G$.
This is not the situation we are in -- we have either commute or coalescence times --
but it is related.
The Random Target Lemma \citep{aldous-fill-2014}
tells us that the stationary distribution of the Markov chain, denoted $\pi$,
can be recovered from the hitting times by solving $\pi = H^{-1} \bone / \bone^T H^{-1} \bone$.
As shown in Appendix \ref{apx::hitting_calcs},
equation \eqref{eqn:hitting_sum} can be rewritten in matrix form as
\begin{align}
    G H = \diag(1/\pi) - \bone \bone^T ,
\end{align}
which implies that $G$ can be computed directly from $H$ as
\begin{align} \label{eqn:G_from_H}
    G = \left( \diag(1/\pi) - \bone \bone^T \right) H^{-1} .
\end{align}
Therefore, $H$ uniquely determines $G$.

@@

% \paragraph{Coalescence times}
The \emph{coalescence times} are defined using \emph{two} copies of the same chain,
as the mean time until coalescence,
if the chains coalesce at rate $\gamma$ when they are in the same place.
Concretely, 
suppose that $X$ and $Y$ are independent Markov chains both moving with movement rates given by $G$,
that coalesce at rate $\gamma_x$ when $X$ and $Y$ are both at $x$.
Then we define $\tau$ to be this coalescence time,
so that 
$\P\{\tau \le t + \epsilon \given \tau > t, \; X_t = Y_t = x\} = \epsilon \gamma_x + O(\epsilon^2)$,
and 
$\P\{\tau \le t + \epsilon \given \tau > t, \; X_t \neq Y_t\} = O(\epsilon^2)$.
Then, the (mean) coalescence time is $C_{xy} = \E[\tau \given X_0 = x, \, Y_0 = y]$.
A similar argument using the Markov property
shows that $C$ satisfies the following equation:
\begin{align}
\text{\bf (coalescence time)} \qquad
    \sum_y \left(G_{xy} C_{yz} + C_{xy} G_{zy}\right)
    &=
    \begin{cases}
        -1                   \quad & \text{if} \; x \neq z, \\
        -1 + \gamma_z C_{zz} \quad & \text{if} \; x = z.
    \end{cases}
\end{align}
Similar recursions are common in the literature,
going back at least to \citet{hill1972effective} (see also \citet{whitlock1997effective}).
In matrix notation, this is
\begin{align} \label{eqn:C_matrix}
    G C + C G^T - \diag(\gamma) \circ C = -\bone \bone^T,
\end{align}
where $\diag(\gamma)$ is the matrix with the vector $\gamma$ on the diagonal and zeros elsewhere,
and $\circ$ is the componentwise product.
In practice, we solve this by working with the product Markov chain $(X_t, Y_t)$,
whose generator matrix is $G \otimes I + I \otimes G$,
where $I$ is the identity matrix and $\otimes$ is the Kronecker product.

@@

This suggests that nonuniqueness of solutions may not be a problem in practice,
as long as geography is discretized into sufficiently many locations
and direct, long-distance migration is disallowed.
However, another problem appears at finer geographic resolutions:
although solutions may be in principle unique,
the problems become \emph{ill-conditioned} 
in the sense that arbitrarily small variations in the input times
(even numerical instability)
may produce large differences in the inferred rates.
Roughly speaking, this occurs because as the spatial resolution increases,
the matrix $G$ converges to a second-order elliptic differential operator \citep{stroock1997multidimensional};
such operators are deformations of the Laplacian \citep{feller4},
which is well-known to have eigenvalues that decay rapidly 
\citep{hpmckean1967curvature,kuttler1984eigenvalues}.
Information about the landscape contained in higher-order modes
(e.g., finer resolution changes in hitting times)
is then almost obliterated by the small eigenvalues of $G$,
making the inverse problem possible in theory but not in practice.
This implies a fundamental limit to the geographical resolution of inferred maps.
For more discussion of related problems, see 
\citet{epstein2008badtruth}, \citet{myers2008learn}, or \citet{terhorst2015fundamental}.

>>>>

\documentclass{article}
\usepackage{fullpage}
\usepackage{amsmath,amssymb,amsthm}
\usepackage[hidelinks]{hyperref}
\usepackage{graphicx}
\usepackage[utf8]{inputenc}
\usepackage{xcolor}
\usepackage{natbib}
% \usepackage{subcaption}
\usepackage{lineno}

@@

\newcommand{\R}{\mathbb{R}}
\newcommand{\norm}[1]{\left\lVert#1\right\rVert}
\DeclareMathOperator{\var}{\mathop{\mbox{Var}}}
\DeclareMathOperator{\cov}{\mathop{\mbox{Cov}}}
\DeclareMathOperator{\diag}{\mathop{\mbox{diag}}}
\renewcommand{\P}{\mathbb{P}}
\newcommand{\E}{\mathbb{E}}
\newcommand{\given}{\,\vert\,}
\newcommand{\supp}{\mathop{\mbox{supp}}}
\newcommand{\sgn}{\mathop{\mbox{sgn}}}
\newcommand{\PP}[1]{\P\!\left\{#1\right\}}
\newcommand{\bone}{\mathbf{1}}

@@

\newif\ifsubmission
% \submissiontrue  % for submission
\submissionfalse % for actual reading

%%% WARNING: right now, with submissiontrue, the supplemental figures dissappear
%%% and are not referred to!!
\ifsubmission
    % \renewcommand{\includegraphics}[2][]{\relax}
    \usepackage{endfloat}
    % \renewcommand{\efloatseparator}{\mbox{}} 
\fi

% Responses to reviews:
% set this to show line numbers and include responses to reviews or not
\newif\ifreviewresponses
\reviewresponsestrue  % include them
% \reviewresponsesfalse  % don't include them
\input{review-response-commands}
\newcommand{\responsefile}{review-responses.tex}  % name of the review reponses file

@@

\linenumbers

@@

Genetic relatedness is determined by past gene flow,
a product of individual or gamete movements across geographic space.
Genomes therefore retain the traces of this movement,
and can contribute to the inference of how a species moves across a landscape,
which is important for understanding how diseases spread, how species adapt,
and how to retain genetic diversity in threatened species.
This is a potentially important source of information,
as direct observation can be difficult 
(especially of long-distance movement, \citep{cayuelademographic,levin_et_al_seed_dispersal}),
or even impossible, if some of the populations in question no longer exist. \revpoint{3}{18}

@@

In this paper, we contrast 
coalescence time and resistance distance,
and develop a method to infer movement rates on a discrete landscape
using genetic distances.
As we will see,
although the two methods are conceptually similar,
resistance-based inference may get the wrong answer if data are derived from a coalescent process,
especially in the presence of gene flow asymmetry. \revpoint{3}{19}
We also explore the question of identifiability.
Resistance-based methods are often used to predict distances based on a few layers of geographical data
(land cover, slope, etcetera).
Some methods find the combination of landscape layers to best explain genetic distance,
which allows very fine geographic resolution \citep{shaffer2017desert}.
Other methods such as EEMS \citep{petkova2016visualizing} infer conductances without such prior information; \revpoint{3}{20}
however, the geographical resolution is much coarser.
(A recently-released method 
that is based on EEMS but uses haplotype length data and an underlying model of coalescence time
\citep{alasadi2018estimating}
is promising, but still produces similarly coarse maps.)
One reason for this difference in resolution
is that EEMS uses more computationally intensive Bayesian methods.
However, we show that this is a more general problem, \revpoint{3}{21}
both because of unmodeled ``process noise'' involved in discretization of geographical populations,
and the mathematical structure of the underlying problem.

@@

There are a number of ways to calculate measures of dissimilarity using genetic data.
In this paper we use mean pairwise divergence, \revpoint{3}{6}
which for two genomes is calculated as the proportion of sites in the genome
at which the two genomes differ. \revpoint{3}{22}
It is more common in the literature to use $F_{ST}/(1-F_{ST})$, \revpoint{1}{7}
but the two are highly correlated for good theoretical reasons 
\citep{slatkin1991inbreeding,rousset1997genetic},
especially for SNP data.
The two genomes differ at a site only if there has been a mutation in some ancestor on the lineages leading
from the two genomes back to their most recent common ancestor at that site.
Therefore,
under an infinite-sites model with average mutation rate $\mu$ per site and per meiosis,
pairwise divergence divided by $\mu$
is an unbiased estimate of the mean time to most recent common ancestor,
averaged across the genome \citep{hudson2007variance,ralph2015empirical}.
It is natural to model this process by following the two lineages back through time
(and across space)
until they find their most recent common ancestor,
i.e., until they coalesce.
Suppose we model the geographic distribution of the species
as a collection of randomly mating populations
that exchange occasional migrants,
thus discretizing geographic space.
If the genome is selectively neutral,
then the motion of these lineages back through time
forms a Markov process, the \emph{structured coalescent} \citep{wakeley2005coalescent}.
In this framework, each lineage performs a random walk across the populations,
with movement probabilities depending on the flux of individuals between populations.
This provides the link between genetic distance and random walks.

@@

Suppose instead we are given the hitting times, $H$, and wish to find the movement rates, $G$.
This is not the situation we are in -- we have either commute or coalescence times --
but it is related.
The Random Target Lemma \citep{aldous-fill-2014}
tells us that the stationary distribution of the Markov chain, denoted $\pi$,
can be recovered from the hitting times by solving $\pi = H^{-1} \bone / \bone^T H^{-1} \bone$.
As shown in Appendix \ref{apx::hitting_calcs},
equation \eqref{eqn:hitting_sum} can be rewritten in matrix form as
\begin{align}
    G H = \diag(1/\pi) - \bone \bone^T ,
\end{align}
which implies that $G$ can be computed directly from $H$ as
\begin{align} \label{eqn:G_from_H}
    G = \left( \diag(1/\pi) - \bone \bone^T \right) H^{-1} .
\end{align}
(We prove that $H$ is invertible in Appendix~\ref{apx::hitting_calcs}.)
Therefore, $H$ uniquely determines $G$.

@@

% \paragraph{Coalescence times}
The \emph{coalescence times} are defined using \emph{two} copies of the same chain,
as the mean time until coalescence,
if the chains coalesce at rate $\gamma$ when they are in the same place.
Concretely, 
suppose that $X$ and $Y$ are independent Markov chains both moving with movement rates given by $G$,
that coalesce at rate $\gamma_x$ when $X$ and $Y$ are both at $x$.
Then we define $\tau$ to be this coalescence time,
so that 
$\P\{\tau \le t + \epsilon \given \tau > t, \; X_t = Y_t = x\} = \epsilon \gamma_x + O(\epsilon^2)$,
and 
$\P\{\tau \le t + \epsilon \given \tau > t, \; X_t \neq Y_t\} = O(\epsilon^2)$.
Then, the (mean) coalescence time is $C_{xy} = \E[\tau \given X_0 = x, \, Y_0 = y]$.
A similar argument using the Markov property
shows that $C$ satisfies the following equation:
\begin{align}
\text{\bf (coalescence time)} \qquad
    \sum_y \left(G_{xy} C_{yz} + C_{xy} G_{zy}\right)
    &=
    \begin{cases}
        -1                   \quad & \text{if} \; x \neq z, \\
        -1 + \gamma_z C_{zz} \quad & \text{if} \; x = z.
    \end{cases}
\end{align}
Similar recursions are common in the literature,
going back at least to \citet{hill1972effective} (see also \citet{whitlock1997effective}).
In matrix notation, this is
\begin{align} \label{eqn:C_matrix}
    G C + C G^T - \diag(\gamma) \circ C = -\bone \bone^T,
\end{align}
where $\diag(\gamma)$ is the matrix with the vector $\gamma$ on the diagonal and zeros elsewhere,
and $\circ$ is the componentwise product.
In practice, we solve this by working with the product Markov chain $(X_t, Y_t)$,
whose generator matrix is $G \otimes I + I \otimes G$, \revpoint{1}{10}
where $I$ is the identity matrix and $\otimes$ is the Kronecker product.

@@

This suggests that nonuniqueness of solutions may not be a problem in practice,
as long as geography is discretized into sufficiently many locations
and direct, long-distance migration is disallowed. \revpoint{3}{2}
However, another problem appears at finer geographic resolutions:
although solutions may be in principle unique,
the problems become \emph{ill-conditioned} 
in the sense that arbitrarily small variations in the input times
(even numerical instability)
may produce large differences in the inferred rates.
Roughly speaking, this occurs because as the spatial resolution increases,
the matrix $G$ converges to a second-order elliptic differential operator \citep{stroock1997multidimensional};
such operators are deformations of the Laplacian \citep{feller4},
which is well-known to have eigenvalues that decay rapidly 
\citep{hpmckean1967curvature,kuttler1984eigenvalues}.
Information about the landscape contained in higher-order modes
(e.g., finer resolution changes in hitting times)
is then almost obliterated by the small eigenvalues of $G$,
making the inverse problem possible in theory but not in practice.
This implies a fundamental limit to the geographical resolution of inferred maps.
For more discussion of related problems, see 
\citet{epstein2008badtruth}, \citet{myers2008learn}, or \citet{terhorst2015fundamental}.
